### 2022-02-13 paper/related_work.tex

In order to estimate normals of an object surface.

In 2012, Holzer et al. \cite{Holzer.S} presented a read-time method, which is able to run algorithm in a high frame speed. They smooth the depth data in order to handle the noise of depth image. The speed is accelerated via integral image. The drawbacks are, as mentioned in the paper, the normals error go up when point depths change severely.

In 2021, Zhou et al.  \cite{zhou2021fast}