### 2022-05-16 paper/introduction.tex
update paper 1. add related works

%% add noise image
%\begin{figure}[!h]
%	\centering
%	\begin{tikzpicture}
%		\node[inner sep=0pt] (depthmap) at (0,0)
%	{\includegraphics[width=.5\textwidth]{./pic/00028.normal0.png}};
%	\end{tikzpicture}
%\end{figure}