### 2022-06-19 paper/Chapters/Chapter06.tex

% Chapter Template

\chapter{Conclusion} % Main chapter title

\label{ch:06} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------
Gated convolution neural network...