### 2022-06-27 paper/Chapters/Chapter02.tex

% Chapter Template

\chapter{Related Work} % Main chapter title

\label{ch:02} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

Guided method \cite{guided} requires addition information like RGB image for the inference.

%% denoising
The depth map captured by depth sensors are usually with full of missing pixels and holes.
\cite{depth-inpainting-distribution} proposed a depth hole filling method using the depth distributions of neighboring regions.

\cite{nconv} introduced normalized convolution dealing with missing or uncertain data for convolution operation. It uses a binary mask to distinguish missing data and integrate it into the convolution operation. \cite{ncnn} applied it and use normalized convolution layers in their networks, which aims to reconstruct the missing pixels from the sparse depth map sensed by cameras. \cite{pncnn} proposed an input confidence estimation network to estimate the confidence instead of using a binary mask. However, the relevant papers are only using 1 channel data as model input, which didn't discussed the case for the multiple-channel data as input.

\cite{gated_activation} proposed a gated activation unit to model more complex interactions comparing to standard CNN layers, which mainly inspired by the multiplicative units exist in Long Short-Term Memory proposed by \cite{lstm} and Rated Recurrent Unit (GRU) proposed by \cite{gru}. 
\cite{gconv} employed the same gated unit solving for the free-form image inpainting task. The proposed network use 3 channel RGB images as input and estimate the missing pixels.

%% method for structured point cloud

%% CNN based methods

estimate the surface normal based on improve the performance of image processing to a brand new stage. 
In 2014, Eigen et al.\cite{Eigen} proposed a method predicting depth map directly from RGB image using CNN. In this case, no depth map is required. In 2016, Laina et al. \cite{img2depth} proposed a deeper network based on ResNet \cite{resnet} with a well designed upsamling part. 
%% guided scene completion

\cite{geometry_based_solution} proposed a method to learn discriminative and geometrically informative primitives from RGBD images, which is further used to recover the surface normals of a scene from a single image.

\cite{GeoNet} uses ResNet proposed by \cite{resnet} to infer the surface normal from RGB image, and further refine it with the help of depth map based on the methods proposed by \cite{geometry_based_solution}  further refine the a  GeoNet to infer depth and surface normal in one unified system,  method proposed by \cite{img2depth}.

%% upsampling 
It is worth to noticed that, the output of normal inference CNN model is not one or severl labels but an entire image or normal map with same size. 
%% talk about image upsampling, unet
Recently, Ronneberger et al proposed an architecture called UNet \cite{unet} for biomedical image segmentations. The architecture is shown in Figure \ref{fig:u-net}.The first half network is a usual classification convolutional network, the second half replace the pooling layers and traditional fc layers in the traditional CNNs to upsampling layers, thus in the end of the second half, the output is able to back to the input size. The proposed network can successfully assigned each pixel a class for segmentation tasks. Under this symmetric network, an input image is downsampled 3 times and upsampled 3 times. Output image has exactly the same size as input image. The downsampling and upsampling both have large number of feature channels, which guarantee the network propagates the information to higher resolution layers.

%% unstructured point cloud
For the unstructured point cloud, the neighbor information of each point is usually unknown. K-nearest neighbor is a common algorithm for neighbor searching. With knowing this information, the neighbor based approaches can be used as a second step. However, KNN-method merely based on the Euclidean distance in the 3D space. Therefore, the points of the other surfaces but in a close distance will also be considered as neighbors. \cite{unstructed-pc} proposed a method based on unstructured point cloud with selecting the optimal scale around each point for normal estimation. \cite{unstructed-pc-patch-stitching} presents a pipeline for calculating the normals of multiple points in parallel.

\section{Standard Methods}

Standard methods compute normals from the point cloud using neighboring information in image space or from a single grayscale image using use Shape from Shading \cite{SFS}. The first method assumes that the neighbors of the points locate on the same plane. This method performs well with a well-chosen window size. However, the drawbacks are that the algorithm is highly noise sensitive. It is weak in handling missing pixels, which is a common issue in the input data. The second method depends on the correct information about the light source. Errors may occur in regions with inter-reflections in the 3D measurement.

%-----------------------------------
%	SUBSECTION 2
%-----------------------------------

\section{Deep Learning based Method}
%% CNN based methods

Recently, deep learning based method achieved a great succeed for image processing.  \cite{yolov3} proposed  \cite{efficientDet} These network architectures use a batch of RGB/Grayscale images as input and are employed for classification problems. Usually, the images are convoluted with a convolution layer and downsampling with pooling layers. The outputs of the network consist of a single value to represent the index of the corresponding class \cite{efficientDet}, or with a set of values to represent the position of bounding boxes.\cite{yolov3}. However, in many other vision tasks, like normal map inference, the output is demanded as the same shape as the input. Instead of predicting one or several classes for the whole input matrix, the class for each pixel requires for prediction. In this case, the traditional network architecture is not suitable anymore.

From PCA for estimation to recently deep learning based methods, the task of surface normal inference is also been well studied in last several decades.

\section{Normal Inference}

%% normal estimation method introduction
Usually, we based on point cloud, depth map or RGB/Grayscale image of the objects or scenes to inference the normals.

Traditional methods evaluate normals based on neighbor information of point cloud or depth map.
In 2012, Holzer et al. \cite{Holzer.S} proposed method to calculate normal from covariance matrices. 
This method use integral image as input, which is able to run algorithm in a high frame speed. They smooth the depth data in order to handle the noise of depth image. The drawbacks are, as mentioned in the paper, the normals error go up when point depths change severely. In 2013, Fouhey et al. \cite{geometry_based_solution} proposed a method constructing a over-determined function systems to predict normals and solving it 
by algebra methods. Similarly, this approach gives a quick but coarse normal inference.