### 2022-08-09 paper/akweury/document.tex

\documentclass[border=15pt, multi, tikz]{article}
\usepackage[backend=bibtex,style=authoryear,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)

\addbibresource{example.bib} % The filename of the bibliography

\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography

\usepackage{amssymb}
\usepackage[backend=bibtex,style=authoryear,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles \left( APA)

\addbibresource{example.bib} % The filename of the bibliography

\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography
\usepackage{import}
\usepackage{float}
\usepackage{tikz}
\usepackage{tikz-network}
\usepackage{breqn}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage{pdfpages}
\usepackage{pgfplots}
\usepackage{rotating}
\usetikzlibrary{fit}
\usetikzlibrary{calc,patterns,angles,quotes}
\usetikzlibrary {arrows.meta,graphs,shapes.misc}
\usetikzlibrary {positioning}

\usetikzlibrary{fit}
\usetikzlibrary {arrows.meta,graphs,shapes.misc}
\usetikzlibrary {positioning}
\subimport{./layers/}{init}
\newcommand{\bn}{\textbf{n}}
\newcommand{\tabhead}[1]{\textbf{#1}}

\def\ConvColor{rgb:yellow,5;red,2.5;white,5}
\def\ConvReluColor{rgb:yellow,5;red,5;white,5}
\def\PoolColor{rgb:red,1;black,0.3}
\def\DcnvColor{rgb:blue,5;green,2.5;white,5}
\def\SoftmaxColor{rgb:magenta,5;black,7}
\def\SumColor{rgb:blue,5;green,15}
\def\poolsep{1}

\begin{document}

\begin{figure}[H]
	\centering

	\begin{tikzpicture}
		\begin{axis} [axis lines=center]
			\addplot [domain=-6:6, thick, smooth, width=5mm] { e^x/(e^x+1) };
		\end{axis}
	\end{tikzpicture}
	\caption{Sigmoid activation function. }
	\label{fig:sigmoid-shape}
\end{figure}

\begin{figure}[H]
	\centering

	\begin{tikzpicture}
		\begin{axis} [axis lines=center]
			\addplot [domain=0:5, thick, smooth, width=5mm] { x };
			\addplot [domain=-5:0, thick, smooth, width=5mm] {0.05*x };
		\end{axis}
	\end{tikzpicture}
	\caption{LeakyRelU activation function. }
	\label{fig:leakyrelu-shape}
\end{figure}

\end{document}